---
title: "Equivalents in R, Python and Perl"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(engine.path = list(perl = '/usr/local/ActivePerl-5.24/bin/perl', python = '/anaconda3/bin/python'))
```

# Introduction

I started a blog post back in 2012 to provide the corresponding code for common programming tasks in [R, Python, and Perl](https://davetang.org/muse/2012/05/03/equivalents-in-r-python-and-perl/). However, at that time I had limited experience with R and Python so the post was very short. I have been using R quite often over the last couple of years and have a better appreciation of the language now.

I have finally decided to learn Python, so I thought I'll revisit my old post and update it. However, it's not convenient to use WordPress. Instead, I'll use R Markdown to document the code since [other languages](https://yihui.name/knitr/demo/engines/) are supported by `knitr`.

# Engines

`knitr` supports various [language engines](https://rmarkdown.rstudio.com/authoring_knitr_engines.html) including Perl and Python. I am using RStudio to edit my R Markdown file and the default Perl and Python interpreter are used to evaluate the code.

```{r default_interpreter}
Sys.which("perl")
Sys.which("python")
```

I don't mind the older Perl version but I would like to use Python 3 and the version in `/usr/bin/` is 2.7. You can specify which interpreter to use by setting `knitr::opts_chunk$set()`; I've set it to use ActivePerl and the Anaconda version of Python. After setting `engine.path`, different interpreters are used.

Perl interpreter version used to evaluate the code in this document.

```{perl}
print("$^X\t$^V\n")
```

Python interpreter version used to evaluate the code in this document.

```{python}
import sys
print(sys.version)
```

## Reticulate

In addition, the [reticulate package](https://github.com/rstudio/reticulate) provides an R interface to Python.

```{r install_reticulate eval = FALSE}
install.packages("reticulate")
```

Use `use_python` to specify the Python interpreter to use.

```{r}
library(reticulate)
use_python("/anaconda3/bin/python")
```

Run some Python code.

```{python}
import sys
print(sys.version)
print(range(0,11))
```

# Hello world

I'll start with a [hello world](https://en.wikipedia.org/wiki/%22Hello,_World!%22_program) example.

In R, `cat` can be used to conceatenate and print out atomic types (those that are composed of only one data type), whereas `print` is more versatile.

```{r r_hello}
cat("Hello world!\n")
print("Hello world!")

# non-atomic data types will produce an erro
# cat(list(a = 1:10))
# cat(data.frame(a = 1:10))

print(list(a = 1:10))
```

In Python, use `print` to print out values.

```{python}
print("Hello world!")
print(1984)
```

In Perl, use `print` to print out values. Notice that I need a newline escape (`\n`) to print the output on separate lines. In addition, I need to terminate each line of code with a semi-colon.

```{perl}
print("Hello world!\n");
print("1984\n")
```

# Data types

A [data type](https://en.wikipedia.org/wiki/Data_type) is a specific classification of data.

In R, there are the character, double, integer, logical, and complex types.

```{r data_type}
typeof("str")
typeof(3.14)
typeof(2L)
typeof(FALSE)
typeof(2i)
```

In Python, there are the string, float, integer, Boolean, and complex types.

```{python}
print(type("str"))
print(type(10.0))
print(type(10))
print(type(False))
print(type(10j))
```

Data structures are also data types in Python, such as lists and dictionaries.

```{python}
print(type([]))
print(type({}))
```

Perl has no type. The following code will execute without the interpreter throwing any errors.

```{perl}
print(1 + 1, "\n");
print(1.0 + 1, "\n");
print(1 + "1", "\n");
print(1e3 + "1", "\n");
print(0xE1 + "1", "\n");
print(1 + "TRUE", "\n");
print("False" + "TRUE")
```

# Data structures

A [data structure](https://en.wikipedia.org/wiki/Data_structure) is a format for storing data. Some structures do not exist in a language, such as lists in Perl.

## Scalars

In Perl, a scalar is a single unit of data. Since Perl has no data types, this can be anything you surround quotes around.

In R, you would store a particular data type into a vector of length one.

```{r scalar}
s <- 'str'
typeof(s)
s <- 1984
typeof(s)
```

In Python, this would be the same.

```{python}
s = 'str'
print(type(s))
s = 1984
print(type(s))
```

In Perl, scalars begin with a dollar sign (`$`).

```{perl}
my $s = 'str';
print("$s\n");
my $s = 1984;
print("$s\n")
```

## Arrays

```{r}
a <- array(1:5)
a

for(i in a){
  print(i)
}
```

In Python.

```{python}
from array import array
a = array('l', [1, 2, 3, 4, 5])

for i in a:
   print(i)

# or use numpy
import numpy
a = numpy.array([1, 2, 3, 4, 5])
b = a/10

print(type(b))

print(b)
```

In Perl:

```{perl}
my @a = (1, 2, 3, 4, 5);
for my $i (@a) {
   print "$i\n";
}
```

## Associative array

In R: https://cran.r-project.org/web/packages/hash/index.html

```{r}
library(hash)
my_hash <- hash(keys = c('colour', 'book', 'movie'),
                values = c('blue', '1984', 'Memento'))

my_hash$movie

my_hash$movie <- 'Inception'
my_hash$movie
del('book', my_hash)

keys(my_hash)
values(my_hash)

my_hash
```

In Python: dictionaries

```{python}
dict = {'colour': 'blue', 'book': '1984', 'movie': 'Memento'}
print(dict['movie'])

dict['movie'] = 'Inception'
print(dict['movie'])

del dict['movie']

print(dict.keys())
print(dict.values())

print(dict)
```

In Perl: hashes

```{perl}
my %hash = ('colour' => 'blue', 'book' => '1984', 'movie' => 'Memento');
print("$hash{movie}\n");

$hash{'movie'} = 'Inception';
print("$hash{movie}\n");

delete($hash{'movie'});

print "$_\n" for keys %hash;
print "$_\n" for values %hash;

print(%hash, "\n")
```

# Loops

## For loops

```{r}
# prints 0 to 9
for (i in 0:9){
  print(i)
}

# using an index
index <- c('one', 'two', 'three')
for (i in index){
  print(i)
}

# example from Pablo
sapply(0:9, FUN = function(x) x )

# load the purrr package
library(purrr)

0:9 %>% map(function(x) x) %>% unlist()
```

```{python}
# prints 0 to 9
for num in range(0, 10):
   print(num)

# using a list
my_list = [1, 2, 3, 4, 5, 6]
for i in my_list:
    print(i)
```

```{perl}
for (my $i = 0; $i < 10; ++$i){
   print "$i\n";
}

for my $i (0 .. 9) {
   print "$i\n";
}
```

## While loops

```{r}
i <- 1
while (i < 10){
  print(i)
  i <- i + 1
}
```

In Python:

```{python}
count = 0
while (count < 9):
   print(count)
   count = count + 1
```

In Perl:

```{perl}
my $i = 0;
while ($i < 10) {
   print ++$i, "\n";
}
```

# Defining functions

In R:

```{r}
multiply_by_two <- function(n) {
   n*2
}
multiply_by_two(2)
```

In Python (see http://docs.python.org/release/1.5.1p1/tut/functions.html):

```{python}
# The keyword def introduces a function definition
def multiply_by_two(n):
   print(n * 2)

multiply_by_two(2)
```

In Perl:

```{perl}
sub multiply_by_two {
   my ($n) = @_;
   #last evaluation is returned by default
   $n*2;
}
print multiply_by_two(2), "\n";
```

# Conditionals

In R:

```{r}
greater_less_equal_5 <- function(answer){
  if (answer > 5){
    print(1)
  }
  else if (answer < 5){
    print(-1)
  } else {
    print(0)
  }
}
greater_less_equal_5(4)
greater_less_equal_5(5)
greater_less_equal_5(6)
```

In Python:

```{python}
def greater_less_equal_5(answer):
    if answer > 5:
        return 1
    elif answer < 5:
        return -1
    else:
        return 0
        
print(greater_less_equal_5(4))
print(greater_less_equal_5(5))
print(greater_less_equal_5(6))
```

In Perl:

```{perl}
sub greater_less_equal_5 {
   $answer = @_[0];
   if ($answer > 5){
      return(1)
   } elsif ($answer < 5){
      return(-1)
   } else {
      return(0)
   }
}

print greater_less_equal_5(4), "\n";
print greater_less_equal_5(5), "\n";
print greater_less_equal_5(6), "\n";
```

# Objects

Firstly, everything in R is treated like as an object. However, R has three object oriented (OO) systems: <a href="http://www.cyclismo.org/tutorial/R/s3Classes.html">[[S3]]</a>, <a href="http://adv-r.had.co.nz/S4.html">[[S4]]</a> and [[R5]]. The basic idea is that a function is defined which creates a list.

```{r}
NorthAmerican <- function(eatsBreakfast=TRUE, myFavorite="cereal"){
   me <- list(hasBreakfast = eatsBreakfast,
              favoriteBreakfast = myFavorite
             )
   ## Set the name for the class
   class(me) <- append(class(me),"NorthAmerican")
   return(me)
}

bubba <- NorthAmerican()
bubba
```

In <a href="https://docs.python.org/2/tutorial/classes.html">Python</a>, the simplest form of class definition looks like this:

```
class ClassName:
    <statement-1>
    .
    .
    .
    <statement-N>

class MyClass:
    """A simple example class"""
    i = 12345
    def f(self):
        return 'hello world'
```

In Perl most objects are hashes and a Perl class is defined in a Perl module file (*.pm) with a package declaration. Here's an <a href="http://perltricks.com/article/25/2013/5/20/Old-School-Object-Oriented-Perl">example</a>:

```{perl}
package Shape;

sub new {
    my $class = shift;
    my $self = {
        color  => 'black',
        length => 1,
        width  => 1,
    };
    return bless $self, $class;
}

1;
```

And a script:

```
#!/usr/bin/env perl

use strict;
use warnings;
use Shape;

# create a new Shape object
my $shape = Shape->new;

# print the shape object attributes
print join("\t", $shape->{color}, $shape->{length}, $shape->{width}), "\n";

exit(0);
```





