---
title: "The reticulate package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The [reticulate](https://rstudio.github.io/reticulate/) package provides a comprehensive set of tools for interoperability between Python and R.

Install the package (if you haven't already).

```{r install}
if(!require("reticulate")){
  install.packages("reticulate")
}
```

Python interpreter in PATH.

```{r which_python_before}
Sys.which("python")
```

Use the Conda Python (`required = TRUE` is necessary to force the change) and print out config.

```{r use_python}
use_python(python = "/Users/dtang/miniconda3/bin/python", required = TRUE)
py_config()
```

Python interpreter after using `use_python`.

```{r which_python_after}
Sys.which("python")
```
Import the Python module `sys` and get Python version.

```{python}
import sys
print(sys.version)
```

To install `numpy`, run `pip install numpy` but make sure you are using the Conda `pip`.

Import `numpy`, create a `numpy` array, and print out `type`.

```{python}
import numpy as np

number = np.array(list(range(11, 17)))
print(type(number))
```
