---
title: "The reticulate package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The [reticulate](https://rstudio.github.io/reticulate/) package provides a comprehensive set of tools for interoperability between Python and R. The package includes facilities for:

* Calling Python from R in a variety of ways including R Markdown, sourcing Python scripts, importing Python modules, and using Python interactively within an R session.
* Translation between R and Python objects (for example, between R and Pandas data frames, or between R matrices and NumPy arrays).
* Flexible binding to different versions of Python including virtual environments and Conda environments.

Reticulate embeds a Python session within your R session, enabling seamless, high-performance interoperability. If you are an R developer that uses Python for some of your work or a member of data science team that uses both languages, reticulate can dramatically streamline your workflow!

Try to load package, installing it if missing.

```{r install}
if(!require("reticulate")){
  install.packages("reticulate")
  require(reticulate)
}
```

Use Miniconda3.

```{r install_miniconda}
if (!dir.exists(miniconda_path())){
  reticulate::install_miniconda()
}
use_python(paste0(miniconda_path(), '/bin/python3'))
```

Python interpreter in PATH.

```{r which_python}
Sys.which("python")
```

Package Installer for Python.

```{r which_pip}
Sys.which("pip")
```

Install `numpy`.

```{bash}
pip install numpy
```

Print out config.

```{r py_config}
py_config()
```

Run some Python code: import the Python module `sys` and get Python version.

```{python}
import sys
print(sys.version)
```

Import `numpy`, create a `numpy` array, and print out `type`.

```{python}
import numpy as np

number = np.array(list(range(11, 17)))
print(type(number))
```
